Analyzing Information Transfer from Different Data Sources via Precise Bias-variance Asymptotics

### 2021-07-02 revision_journal/exp.tex
restructure
reorg wip
update intro wip
Updates from Overleaf

\section{Theoretical Implications}

%\subsection{Comparison between the averaging and HPS estimators}

\subsection{Covariance Alignment}

\paragraph{Covariate shift}
Recall from Example \ref{ex_covshift} that having covariate shifts worsens the variance (hence the loss) of hard parameter sharing when the sample ratio increases.
This highlights the need for correcting covariate shifts when the sample size ratio rises.
To this end, we study a covariance alignment procedure proposed in \cite{WZR20}, designed to correct covariate shifts.
The idea is to add an alignment module between the input and the shared module $B$.
This module is then trained together with $B$ and the output layers. We refer to \cite{WZR20} for more details about the procedure and the implementation.
%We validate our insight on this procedure in the experiments.
%We implement the covariance alignment procedure following \cite{WZR20}.

@@

\subsection{Progressive Training}%\label{sec_text}

Our results and simulations are all in the high-dimensional linear regression setting.
How well do they extend to other scenarios?
In this section, we conduct further studies on six text classification datasets.
Our datasets include a movie review sentiment dataset (MR) \cite{pang2005seeing}, a sentence subjectivity dataset (SUBJ) \cite{pang2004sentimental}, a customer reviews dataset (CR) \cite{hu2004mining}, a question type dataset (TREC) \cite{li2002learning}, an opinion polarity dataset (MPQA) \cite{wiebe2005annotating}, and the Stanford sentiment treebank (SST) dataset \cite{socher2013recursive}.
%The question is to predict positive or negative sentiment expressed in the text.
Our model consists of a word embedding layer with GloVe embeddings \cite{pennington2014glove} followed by a long-short term memory (LSTM) or a multi-layer perception (MLP) layer \cite{lei2018simple}.\footnote{For MLP, we apply an average pooling layer over word embeddings. For LSTM, we add a shared feature representation layer on top of word embeddings.}

%multi-layer perceptron (MLP), LSTM, CNN on all tasks
%We use this task to verify our theoretical results on model capacity and task covariance in real world.
%{\it ChestX-ray14.} This dataset contains 112,120 frontal-view X-ray images \cite{chexnet17}.
%There are 14 diseases (tasks) for every image that we would like to predict.
%We use densenet121 as the shared module \cite{huang2017densely}.
%We treat each label as one task a binary classification problem and formulate it as a 14-task multi-task learning problem.
%This dataset is curated where the labels
%We use the CheXNet model from~\cite{chexnet}, which is a 121-layer convolutional neural network on all tasks.
%For the text classification experiment, we encode each word using the GLoVe word embeddings.%
%\footnote{http://nlp.stanford.edu/data/wordvecs/glove.6B.zip}
%We evaluate three model choices.
%\textit{Predicting transfer effect via STL results.}
%We show that the single-task based metric proposed in Section \ref{sec_similarity} can predict positive or negative transfer in MTL.
%A common challenge in the study of MTL is that the results can be hard to understand.
%It is difficult to predict when MTL performs well without running extensive trials.
%Our insight is that we can use STL results to help understand MTL results.
%Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 tasks.
%We find that using a threshold of $\tau = 0.1$, the STL results correctly predict positive or negative transfer with $75.6\%$ precision and $38.8\%$ recall among $30$ times $5$ (random seeds) task pairs!
%We observe similar results for $91$ task pairs from the ChestX-ray14 dataset.
%The results show that STL results are indicative of MTL results.

\paragraph{Sample size ratio.}
First, we we show that our observation in Figure \ref{fig_size} also occurs in the text classification tasks.
In Figure \ref{fig_ab_data}, we observe that for multiple example task pairs, increasing task one's sample size improves task two's prediction accuracy initially, but hurts eventually.
On the $y$-axis, we plot task two's test accuracy using HPS, subtracted by its STL test accuracy.
%validate that as we increase the sample ratio while keeping task two's sample size fixed, task two's prediction accuracy does not always increase.
We fix task two's sample size at $1000$ and increase task one's sample size from $100$ to $3000$.

\subsection{Rank Selection}
%As a further validation, excluding TREC, we observe similar comparative results.
%The data efficiency ratio of using MLP is $100\%$ because the average performance of MTL is worse than the average of STL.
%We further show that applying incremental training helps reduce the data efficiency ratio to \alert{$xx\%$}.
%If TREC is not included, we see that only $25\%$ of the labeled data is needed.

>>>>

\section{Empirical results}

%\subsection{Comparison between the averaging and HPS estimators}
\subsection{Regression tasks}

\subsection{Text classification tasks}
Our results and simulations are all in the high-dimensional linear regression setting.
How well do they extend to other scenarios?
In this section, we conduct further studies on six text classification datasets.
Our datasets include a movie review sentiment dataset (MR) \cite{pang2005seeing}, a sentence subjectivity dataset (SUBJ) \cite{pang2004sentimental}, a customer reviews dataset (CR) \cite{hu2004mining}, a question type dataset (TREC) \cite{li2002learning}, an opinion polarity dataset (MPQA) \cite{wiebe2005annotating}, and the Stanford sentiment treebank (SST) dataset \cite{socher2013recursive}.
%The question is to predict positive or negative sentiment expressed in the text.
Our model consists of a word embedding layer with GloVe embeddings \cite{pennington2014glove} followed by a long-short term memory (LSTM) or a multi-layer perception (MLP) layer \cite{lei2018simple}.\footnote{For MLP, we apply an average pooling layer over word embeddings.
For LSTM, we add a shared feature representation layer on top of word embeddings.}

\paragraph{Progressively increasing source dataset sizes.}%\label{sec_text}
%multi-layer perceptron (MLP), LSTM, CNN on all tasks
%We use this task to verify our theoretical results on model capacity and task covariance in real world.
%{\it ChestX-ray14.} This dataset contains 112,120 frontal-view X-ray images \cite{chexnet17}.
%There are 14 diseases (tasks) for every image that we would like to predict.
%We use densenet121 as the shared module \cite{huang2017densely}.
%We treat each label as one task a binary classification problem and formulate it as a 14-task multi-task learning problem.
%This dataset is curated where the labels
%We use the CheXNet model from~\cite{chexnet}, which is a 121-layer convolutional neural network on all tasks.
%For the text classification experiment, we encode each word using the GLoVe word embeddings.%
%\footnote{http://nlp.stanford.edu/data/wordvecs/glove.6B.zip}
%We evaluate three model choices.
%\textit{Predicting transfer effect via STL results.}
%We show that the single-task based metric proposed in Section \ref{sec_similarity} can predict positive or negative transfer in MTL.
%A common challenge in the study of MTL is that the results can be hard to understand.
%It is difficult to predict when MTL performs well without running extensive trials.
%Our insight is that we can use STL results to help understand MTL results.
%Table \ref{tab:mtl_better_than_stl} shows the result on both the sentiment analysis and the ChestX-ray14 tasks.
%We find that using a threshold of $\tau = 0.1$, the STL results correctly predict positive or negative transfer with $75.6\%$ precision and $38.8\%$ recall among $30$ times $5$ (random seeds) task pairs!
%We observe similar results for $91$ task pairs from the ChestX-ray14 dataset.
%The results show that STL results are indicative of MTL results.
%\paragraph{Sample size ratio.}
First, we we show that our observation in Figure \ref{fig_size} also occurs in the text classification tasks.
In Figure \ref{fig_ab_data}, we observe that for multiple example task pairs, increasing task one's sample size improves task two's prediction accuracy initially, but hurts eventually.
On the $y$-axis, we plot task two's test accuracy using HPS, subtracted by its STL test accuracy.
%validate that as we increase the sample ratio while keeping task two's sample size fixed, task two's prediction accuracy does not always increase.
We fix task two's sample size at $1000$ and increase task one's sample size from $100$ to $3000$.

\paragraph{Covariate alignment.}
Recall from Example \ref{ex_covshift} that having covariate shifts worsens the variance (hence the loss) of hard parameter sharing when the sample ratio increases.
This highlights the need for correcting covariate shifts when the sample size ratio rises.
To this end, we study a covariance alignment procedure proposed in \cite{WZR20}, designed to correct covariate shifts.
The idea is to add an alignment module between the input and the shared module $B$.
This module is then trained together with $B$ and the output layers. We refer to \cite{WZR20} for more details about the procedure and the implementation.
%We validate our insight on this procedure in the experiments.
%We implement the covariance alignment procedure following \cite{WZR20}.

@@

%\subsection{Model width selection}
%As a further validation, excluding TREC, we observe similar comparative results.
%The data efficiency ratio of using MLP is $100\%$ because the average performance of MTL is worse than the average of STL.
%We further show that applying incremental training helps reduce the data efficiency ratio to \alert{$xx\%$}.
%If TREC is not included, we see that only $25\%$ of the labeled data is needed.

### 2021-07-02 revision_journal/intro.tex
Updates from Overleaf

\paragraph{Random matrix theory.}
\todo{write a paragraph regarding applications of RMT to modern machine learning.}
\begin{itemize}
	\item \citet{hastie2019surprises}
	\item \citet{bartlett2020benign}
	\item \citet{liang2020just}
	\item \citet{montanari2019generalization}, \citet{liang2020precise}
\end{itemize}

Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, a recent development in the random matrix theory literature.
In the single-task case, \cite{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
\cite{Anisotropic} later extended this result to sample covariance matrices with arbitrary covariance.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
These techniques provide almost sharp convergence rates to the asymptotic limit compared to other methods such as free probability \cite{nica2006lectures}.
To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for two tasks (with arbitrary covariate shifts).
%\HZ{can you please add a discussion of the related work from RMT?}

@@

>>>>

\paragraph{Random matrix theory.}
\todo{write a paragraph regarding applications of RMT to modern machine learning.}
\begin{itemize}
	\item \citet{hastie2019surprises}
	\item \citet{bartlett2020benign}
	\item \citet{liang2020just}
	\item \citet{montanari2019generalization}, \citet{liang2020precise}
\end{itemize}
\begin{remark}
    \todo{clarify overparametrize}
    As a clar,  following the work of \citet{hastie2019surprises} (see also \citet{bartlett2020benign} and the references therein)
\end{remark}

In the single task case, if the data is Gaussian distributed, then precise variance can be derived immediately from properties of Wishart distribution. In particular, in the high dimensional setting, the eigenvalues of Wishart matrices satisfy the famous Marchenko–Pastur (MP) law \cite{MP}. It well-known that the MP law holds universally regardless of the detailed data distribution \cite{bai2010spectral}. 
In the single-task case, \cite{isotropic} first proved such a local law for sample covariance matrices with isotropic covariance.
\cite{Anisotropic} later extended this result to sample covariance matrices with arbitrary covariance.
%On the other hand, one may derive the asymptotic result in Theorem \ref{thm_main_RMT} with error $\oo(1)$ using the free addition of two independent random matrices in  theory .
Our proof techniques use the so-called local law of random matrices \cite{erdos2017dynamical}, a recent development in the random matrix theory literature.
These techniques provide almost sharp convergence rates to the asymptotic limit compared to other methods such as free probability \cite{nica2006lectures}.
To the best of our knowledge, we are not aware of any previous results in the multi-task case, even for two tasks (with arbitrary covariate shifts).
%\HZ{can you please add a discussion of the related work from RMT?}

@@

%\subsection{Notations}

%for a small constant $\tau>0$.
